Word learning at the semantics pragmatics interface

### 2020-04-16 paper/supplementary_material.Rmd
further details of functional form of rsa components

$$
P_{L}(r \mid u)\propto P_{S}(u \mid r) \cdot P(r \mid \rho_i) \\ 
P_{S}(u \mid r)\propto Informativity(u;r)^{\alpha_i} \\
Informativity(u; r) = P(r|u) \propto P(r \mid \rho_i) \cdot \mathcal{L}(u \mid \theta_{ij})
$$

@@

The informativity of each utterance is given by imagining which referent a literal listener, who interprets words according to their lexicon $\mathcal{L}$, would infer upon hearing the utterance. Thus, this reasoning depends on what kind of semantic knowledge (word--object mappings) the speaker thinks the literal listener knows. We parameterize the  listener's knowledge of a word's semantics in terms of a semantic knowledge parameter $\theta$, which varies between 0 and 1. $\theta = 0$ correspondes to the state of knowledge for a completely novel word and results in a semantic interpretation function that chooses randomly between the objects in the scene. Each of the novel words are assumed to have semantic knowledge of 0. For $\theta \in (0, 1)$, the semantic interpretation function will select the familiar referent with probability $\theta + (1 - \theta)\frac{1}{2} = \frac{1 + \theta}{2}$; that is, with probability $\theta$, the listener knows the correct meaning of the word (and picks out the correct referent 100\% of the time); with probability $1 - \theta$, the listener does not know the meaning of the word and must guess, picking out the correct referent 50\% of the time.^[
  One could also define $\theta$ to directly denote the probability of selecting the familiar~vs.~unfamiliar referent. The model behavior would be the same except for a translation of this parameter. We choose the parametrization described because we find it more intuitive for $\theta$ to represent the degree of semantic knowledge, where 0 semantic knowledge represents a state of complete uncertainty.
] For familiar objects, semantic knowledge is a function of the degree-of-acquisition of the associated word, which in turn depends upon the word $j$ (its expected acquisition trajectory) as well as on the child's age $i$.

@@

Sensitivity to common ground $\rho$ refers to the probability that an object is taken to be the referent of the utterance before actually hearing the utterance. Thus, it captures the salience of an object due to its role in the social interaction that precedes the utterance. We expect children at different ages to respond differently to the common ground manipulation, resulting in an age specific prior distribution over objects [akhtar1996role; @diesendruck2004two], which we model with a logistic-linear function: $\rho_i = \text{logistic}(\beta^\rho_0 + i \cdot \beta^\rho_1)$.

@@

Next, we combined the parameters according to the four models described below. Note that the parameter distributions were the same for all models (see Figure \@ref(fig:fig4)) and that models only differed in terms of which parameters they included. The models described below are a full model (*integration model*) and three lesioned models, which selectively omit one type of information. The following model comparison therefore asks which types of information are necessary to make good predictions about how information is integrated. We do not compare models that make different assumptions about how information is integrated, since they require additional parameters specific to Experiment 3. We consider the question of alternative integration models in the [explanation](#explanation) section.

@@

The integration model serves as the full model and takes in all available information. That is, it takes in object-specific semantic knowledge, speaker informativeness and common ground sensitivity and combines these components by way of the process described above. Figure \@ref(fig:fig5) visualizes the corresponding model predictions in comparison to the data from experiment 3.

@@

The first lesionsed model takes in speaker informativeness and common ground sensitivity as well as general semantic knowledge, but omits semantic knowledge that is specific to the familiar objects. We described above that the parameters for semantic knowledge are fitted via a hierarchical regression (mixed effects) model. In this model, there is an overall developmental trajectory for semantic knowledge (main effect) and then there is object-specific variation around this trajectory (random effects). The *no word knowledge* model takes in the overall trajectory for semantic knowledge but ignores object-specific variation. That is, the model assumes a listener whose mutual exclusivity inference does not vary depending on the particular familiar object but only depends on the average semantic knowledge.

@@

This model takes in object specific semantic knowledge and speaker informativeness but ignores common ground. Thus, the prior distribution over objects in the model described above is uniform (e.g., [0.5, 0.5]). This corresponds to a listener who only focuses on the mutual exclusivity inference and ignores the common ground manipulation. As a consequence, the listener does not differentiate between the two common ground alignment conditions.

>>>>

$$
P_{L}(r \mid u)\propto P_{S}(u \mid r) \cdot P(r \mid \rho_i) \\ 
P_{S}(u \mid r)\propto Informativity(u;r)^{\alpha_i} \\
Informativity(u; r) = P(r|u) \propto \mathcal{L}(u \mid \theta_{ij})
$$

@@

The informativity of each utterance is given by imagining which referent a literal listener, who interprets words according to their lexicon $\mathcal{L}$, would infer upon hearing the utterance.^[
  Following @Frank2014, we use an implicit uniform prior over referents in the literal listener.
] Thus, this reasoning depends on what kind of semantic knowledge (word--object mappings) the speaker thinks the literal listener knows. We parameterize the  listener's knowledge of a word's semantics in terms of a semantic knowledge parameter $\theta$, which varies between 0 and 1. $\theta = 0$ correspondes to the state of knowledge for a completely novel word and results in a semantic interpretation function that chooses randomly between the objects in the scene. Each of the novel words are assumed to have semantic knowledge of 0. For $\theta \in (0, 1)$, the semantic interpretation function will select the familiar referent with probability $\theta + (1 - \theta)\frac{1}{2} = \frac{1 + \theta}{2}$; that is, with probability $\theta$, the listener knows the correct meaning of the word (and picks out the correct referent 100\% of the time); with probability $1 - \theta$, the listener does not know the meaning of the word and must guess, picking out the correct referent 50\% of the time.^[
  One could also define $\theta$ to directly denote the probability of selecting the familiar~vs.~unfamiliar referent. The model behavior would be the same except for a translation of this parameter. We choose the parametrization described because we find it more intuitive for $\theta$ to represent the degree of semantic knowledge, where 0 semantic knowledge represents a state of complete uncertainty.
] For familiar objects, semantic knowledge is a function of the degree-of-acquisition of the associated word, which in turn depends upon the word $j$ (its expected acquisition trajectory) as well as on the child's age $i$.

@@

Sensitivity to common ground $\rho$ refers to the probability that an object is taken to be the referent of the utterance, before actually hearing the utterance. In our paradigm, common ground takes the form of *discourse novelty*, and $\rho$ captures the salience (or lack thereof) of an object due to its presence or absence in the social interaction that precedes the utterance. Specifically, the pragmatic listener uses the common ground parameter $\rho$ as the prior probability of the discourse-novel (i.e., the referent that was absent in the social interaction that preceded the utterance): In the *congruent* condition, the object in the prior social interaction (i.e., the discourse-familiar object) was the familiar object, and hence, both common ground (discourse novelty) and mutual exclusivity are cues to the same referent; in the *incongruent* condition, the object in the prior social interaction was the novel object, and hence, common ground (discourse-novelty) and mutual exclusitivity are cues to different referents.  
We expect children at different ages to respond differently to the common ground manipulation, resulting in an age specific prior distribution over objects [akhtar1996role; @diesendruck2004two], which we model with a logistic-linear function: $\rho_i = \text{logistic}(\beta^\rho_0 + i \cdot \beta^\rho_1)$.

@@

Next, we combined the parameters according to the four models described below. Note that the parameter distributions were the same for all models (see Figure \@ref(fig:fig4)) and that models only differed in terms of which parameters they included. The models described below are a full model (*integration model*) and three lesioned models, which selectively omit one type of information. The following model comparison therefore asks which types of information are necessary to make good predictions about how information is integrated. All models used age-specific parameters when the parameter was present. We do not compare models that make different assumptions about how information is integrated, since they require additional parameters specific to Experiment 3. We consider the question of alternative integration models in the [explanation](#explanation) section.

@@

The integration model serves as the full model and takes in all available information. That is, it takes in object-specific semantic knowledge, speaker informativeness and common ground sensitivity and combines these components by way of the process described above. Figure \@ref(fig:fig5) visualizes the corresponding model predictions in comparison to the data from Experiment 3.

@@

The first lesioned model takes in speaker informativeness and common ground sensitivity as well as general semantic knowledge (vocabulary development), but omits semantic knowledge that is specific to the familiar objects. We described above that the parameters for semantic knowledge are fitted via a hierarchical regression (mixed effects) model. In this model, there is an overall developmental trajectory for semantic knowledge (main effect) and then there is object-specific variation around this trajectory (random effects). The *no word knowledge* model takes in the overall trajectory for semantic knowledge---represented by $\mu^\theta_0$ and $\mu^\theta_1$---but ignores object-specific variation. Formally, the semantic knowledge for each item is the same: $\theta_{ij} = \mu^\theta_0 +  i \cdot \mu^\theta_1$. That is, the model assumes a listener whose mutual exclusivity inference does not vary depending on the particular familiar object but only depends on the average semantic knowledge.

@@

This model takes in object specific semantic knowledge and speaker informativeness but ignores common ground $\rho_i$. Thus, the prior distribution over objects in the model described above is uniform (e.g., [0.5, 0.5]). This corresponds to a listener who only focuses on the mutual exclusivity inference and ignores the common ground manipulation. As a consequence, the listener does not differentiate between the two common ground alignment conditions.
